---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Creating interactive web maps from raster datasets in R

To run this code you'll need the following R packages installed:

```{r}
library(sf)
library(tmap)
library(osmdata)
tmap_mode("view")
```


## Data cleaning / preprocessing

The first stage is to check the size of datasets in your project.

```{r}
ip = "\\.jpg|\\.JPG|\\.tif|\\.TIF"
f = list.files(path = ".", pattern = ip)
f_size = file.size(f) / 1e6
f_df = data.frame(f, f_size, stringsAsFactors = FALSE)
f_df = f_df[order(f_size, decreasing = TRUE), ]
# head(f_df)
```

```{r, echo=FALSE}
knitr::kable(head(f_df), digits = 0)
```

Some of these are huge files.
That is fine if time and computing resources including RAM and internet bandwidth etc are not a problem limitation.
But for the purposes of testing and teaching methods, it makes sense to compress the images before the other steps.

First we can do this by changing the resolution.
In the next line of code, we will set the maximum number of pixels, in X or Y dimensions, to 1000 pixels.
This parameter will be used subsequently in the code.

```{r}
px_max = 1000 # other max pixel values can be used here
```

To compress images from the command line you can use ImageMagick.
We'll do this for the 5th largest image in the dataset, as follows:

```{r}
j = 5
i = magick::image_read(f_df$f[j])
magick::image_info(i)
ic = magick::image_resize(i, geometry = paste0(px_max, "x", px_max))
magick::image_info(ic)
magick::image_write(ic, "c-image-test.tif")
file.size(f_df$f[j]) / 1e6
file.size("c-image-test.tif") / 1e6
```

What just happened? 
We just reduced a 10 MB file to a 0.1 MB file, but reducing the resolution.
Now we can do this for all datasets (not run as this is slow):

```{r}
dir.create("c-res")
f_df$file_name_res = paste0("c-res/", f_df$f)
f_df$f_size_res = NA
# Note: uncomment the 3 commented lines below to run the compression code
for(j in 1:nrow(f_df)) {
  # i = magick::image_read(f_df$f[j])
  # ic = magick::image_resize(i, geometry = "1000x1000")
  # magick::image_write(ic, f_df$file_name_res[j])
  f_df$f_size_res[j] = file.size(f_df$file_name_res[j]) / 1e6
}
```

```{r}
sum(f_df$f_size)
sum(f_df$f_size_res, na.rm = TRUE)
sum(f_df$f_size) / sum(f_df$f_size_res, na.rm = TRUE)
plot(f_df$f_size, f_df$f_size_res)
knitr::kable(head(f_df[c(1, 2, 4)]), digits = 0)
```

Overall the code above has reduced the image data by more than 1 GB, with the a maximum compressed file size of around 3 MB.

Another way to reduce the size of images is by saving them in a compressed file format (.tif files can be compressed with GDAL, in this case we will save them as .jpg files with the default quality settings).

```{r}
dir.create("c-jpg")
f_df$file_name_jpg = gsub(pattern = ip, ".jpg", x = f_df$f)
f_df$file_name_jpg = paste0("c-jpg/", f_df$file_name_jpg)
f_df$f_size_jpg = NA
# Note: uncomment the 2 commented lines below to run the compression code
for(j in 1:nrow(f_df)) {
  # i = magick::image_read(f_df$f[j])
  # magick::image_write(i, f_df$file_name_jpg[j], quality = 30, format = "jpeg")
  f_df$f_size_jpg[j] = file.size(f_df$file_name_jpg[j]) / 1e6
}
```

```{r}
sum(f_df$f_size_jpg)
sum(f_df$f_size) / sum(f_df$f_size_jpg)
plot(f_df$f_size, f_df$f_size_jpg)
knitr::kable(head(f_df[c(1, 2, 6)]), digits = 0)
```

This shows that saving the files in a compressed format, `.jpeg` with a high compression ratio (quality can be increased by changing 30 to a higher value, up to 100) has also reduced image sizes by more than 1GB.

We can use the raw input files, the lower resolution files, or the compressed .jpeg files.
With computational operations, it is often sensible to try a method first on a small dataset and only after tweaking the methods/settings on things that run quickly, run it on the whole dataset.

## Vector data 

Vector data is available from a wide range of places.
OpenStreetMap is a good up-to-date source that has global coverage and is ever improving, albeit with uneven quality, thanks to thousands of volunteers.

```{r}
library(osmdata)
osm_data = opq(bbox = "hereford uk") %>% 
  add_osm_feature("name", "Rotherwas Industrial Estate") %>% 
  osmdata_sf()
rotherwas_boundary = osm_data$osm_polygons
class(rotherwas_boundary)
plot(rotherwas_boundary)
```

```{r, echo=FALSE, eval=FALSE}
saveRDS(rotherwas_boundary, "rotherwas_boundary.Rds")
```

We can plot this in an interactive mapping environment as follows:

```{r}
tm_shape(rotherwas_boundary) +
  tm_borders()
```


